Evolution of the AGN UV luminosity function from redshift 7.5

### 2018-05-19 Paper/paper.tex
Update text
Comments from Joe

\begin{abstract}
  Determinations of the UV luminosity function of AGN at high
  redshifts are important for constraining the contribution of AGN to
  reionization.  Many recent inferences of the luminosity function
  suffer from inconsistencies arising from inhomogeneous selection and
  mixing of AGN data.  We address this problem in this paper by
  constructing a sample of more than 85,000 colour-selected AGN from
  redshift $z=0$ to $7.5$.  This sample is constructed out of multiple
  data sets with spectroscopic redshifts and completeness estimates.
  We homogenise these data sets to identical cosmologies, intrinsic
  AGN spectra, and magnitude systems.  Using this sample, we derive
  the AGN UV luminosity function from redshifts $z=0$ to $7.5$.  The
  luminosity function has a double power law form at $z<5.5$, and
  single power law form at higher redshifts.  The break luminosity
  brightens rapidly with redshift and the faint end of the luminosity
  function steepens.  In spite of this steepening, the contribution of
  AGN to the hydrogen photoionization rate at $z=5$--$6$ is
  subdominant, although it can be non-negligible if these luminosity
  functions hold down to $M_{1450}=-18$.  Under reasonable
  assumptions, AGN can reionize \HeII\ by redshift $z=3.2$.  At low
  redshifts ($z<0.5$), AGN can produce all of the hydrogen
  photoionization rate implied by the \lya data.  Our global analysis
  of the luminosity function also reveals important systematic errors
  in the data, particularly at $z<0.6$ and $z=2.2$--$3.5$, which need
  to be addressed in future in order to improve our results.  We make
  various fitting functions, luminosity function analysis codes, and
  homogenised AGN data publicly available.
\end{abstract}

@@

The luminosity function of active galactic nuclei (AGN) and its
evolution over cosmological time scales has been a matter of central
interest of a large body of work over the last four decades
\citep[e.g.,][]{1978A&A....68...17M, 1983ApJ...269..352S,
  1988ApJ...325...92K, 1988MNRAS.235..935B, 1993ApJ...406L..43H,
  1994ApJ...421..412W, 1995AJ....110...68S, 1995AJ....110.2553K,
  1995ApJ...438..623P, 2000MNRAS.317.1014B, 2001AJ....121...54F,
  2004AJ....128..515F, 2006AJ....131.2766R, 2007ApJ...654..731H,
  2009MNRAS.392...19C, 2010AJ....139..906W, 2011ApJ...728L..26G,
  2013ApJ...773...14R, 2013ApJ...768..105M, 2015AA...578A..83G,
  2015ApJ...798...28K, 2016ApJ...829...33Y, 2016ApJ...833..222J,
  2017MNRAS.466.1160M}.  Determination of the AGN luminosity function
constrains models of the build-up of supermassive black holes
\citep{2015MNRAS.452..575S, 2016MNRAS.462..190R}.  Due to the
incidence of supermassive black holes in most galaxies, the tight
scaling relations observed to exist between the mass of these black
holes and properties of their host galaxies
\citep{2013ARA&A..51..511K, 2013ApJ...764..184M}, and the increasing
consensus that AGN activity feeds back on the host galaxy evolution,
the AGN luminosity function also constrains models of galaxy
formation.  Finally, thanks to their relative brightness and high
Lyman-continuum (LyC) escape fractions, the luminosity function of AGN
determines their influence on the temperature and ionization state of
the intergalactic medium (IGM) over large scales, possibly even
primarily driving hydrogen and helium reionization.

There is a renewed interest in understanding the evolution of the UV
luminosity function of AGN, triggered by the 19 low-luminosity
($M_{1450}>-22.6$) AGN candidates between redshifts $z=4.1$ and $6.3$
discovered by \citet{2015AA...578A..83G} using a novel X-ray/NIR
selection criterion.  This finding suggested that AGN brighter than
$M_{1450}=-18$ can potentially produce all of the metagalactic
hydrogen photoionization rate inferred from the \lya forest at
$4<z<6$.  A significant presence of AGN at high redshift ($z\sim 6$)
and a dominant contribution of AGN to reionization is appealing as the
LyC escape fraction of galaxies is uncertain.  High-redshift galaxies
down to rest-frame UV magnitude $M_\mathrm{UV}=-12.5$ ($L\sim
10^{-3}L^*$) at $z=6$ \citep{2017ApJ...835..113L} and redshifts up to
$z=11.1$ \citep{2016ApJ...819..129O} have now been reported.  But the
escape of LyC photons has been measured in only a few comparatively
bright ($L>0.5L^*$) galaxies at relatively low redshifts ($z < 4$).
The escape fraction in these galaxies is inferred to be typically
about 2--20\% \citep{2010ApJ...725.1011V, 2011ApJ...736...41B,
  2015ApJ...804...17S, 2015ApJ...810..107M, 2016A&A...585A..48G,
  2017MNRAS.468..389J, 2017MNRAS.465..316M}, whereas reionization
requires escape fraction of about 20\% in galaxies as faint as
$M_\mathrm{UV}=-13$ \citep{2016PASA...33...37F, 2015ApJ...802L..19R,
  2016MNRAS.457.4051K}.  An enhanced incidence of high-redshift AGN
may also be consistent with the shallow bright-end slopes of the
$z\sim 7$ UV luminosity function of galaxies relative to a Schechter
function \citep{2012MNRAS.426.2772B, 2014MNRAS.440.2810B,
  2014ApJ...792...76B, 2015MNRAS.452.1817B} and the hard spectra of
these galaxies \citep{2015MNRAS.450.1846S, 2015MNRAS.454.1393S,
  2017MNRAS.464..469S}.  Finally, AGN may also provide a natural
explanation for the large scatter in the \lya opacity between
different quasar sightlines close to redshift $z=6$
\citep{2015MNRAS.447.3402B, 2018arXiv180208177B}.  These opacity
fluctuations extend to substantially larger scales ($\gtrsim 50\,
h^{-1}$cMpc) than expected in galaxies-dominated reionization models
\citep{2015MNRAS.453.2943C}.  AGN clustering can result in these
fluctuations naturally if there is a significant contribution
($\gtrsim 50\%$) of AGN to the ionising emissivity at $z=5$--$6$
\citep{2017MNRAS.465.3429C}.

Several counter-arguements against a higher incidence of AGN at high
redshift have also been made. \citet{2017MNRAS.468.4691D} and
\citet{2018MNRAS.473.1416M} pointed out that an AGN population
consistent with the findings of \citet{2015AA...578A..83G} leads to a
much earlier \HeII\ reionization.  For instance, in the model of
\citet{2015ApJ...813L...8M}, He~\textsc{ii} reionization is complete
at $z=4.5$, compared to $z=3$ in the standard scenario
\citep{2012ApJ...746..125H}.  Such early \HeII\ reionization could
result in higher IGM temperatures due to the associated photoheating.
\citet{2017MNRAS.468.4691D} found that the temperature of the IGM at
mean density is twice as much in AGN-dominated reionization models as
the standard models at $z=3.5$--$5$, in conflict with constraints from
the \lya forest.  This inconsistency could be avoided by postulating a
reduced escape fraction of \HeII-ionizing photons in AGN, but it is
difficult to reconcile this with a unit escape fraction of
hydrogen-ionizing photons that is required to explain the Ly$\alpha$
opacity fluctuations.  Further arguments against AGN-dominated
reionization have been presented by \citet{2016MNRAS.459.2299F} by
analysing the incidence of metal-line absorbers at $z\sim 6$.
\citet{2016MNRAS.459.2299F} find that in their cosmological radiation
hydrodynamical simulations AGN-dominated UV background results in too
many C~\textsc{iv} absorption systems relative to Si~\textsc{iv} and
C~\textsc{ii} at $z\sim 6$.  However, these simulations assume a
$L_\nu\propto\nu^{-1.57}$ AGN SED at extreme UV
\citep{2002ApJ...565..773T}, which somewhat harder than recent
measurements ($L_\nu\propto\nu^{-1.7}$) by
\citet{2015MNRAS.449.4204L}.  \citet{2016MNRAS.459.2299F} also found
that the N(Si~\textsc{iv})/N(C~\textsc{iv}) column density ratio
measurements prefer a harder and more intense $>4$~Ry background than
the standard model \citep{2012ApJ...746..125H}.  Finally, comparing
the \citet{2015AA...578A..83G} sample to X-ray-selected quasar data at
$z=0$--$6$, \citet{2017MNRAS.465.1915R} argued that the faint end of
the AGN UV luminosity function at $z\sim 6$ is probably shallower that
that reported by \citet{2015AA...578A..83G}.
\citet{2017MNRAS.465.1915R} argue that the apparent contradiction with
the results of \citet{2015AA...578A..83G} could be explained by
contamination from the AGN host galaxies for faint AGN.

@@

\begin{figure*}
  \begin{center}
    \includegraphics[width=\textwidth]{qsos.pdf}
    % data.py 
  \end{center}
  \caption{Redshift distribution of the 83,491 AGN considered in this
    analysis.  Shown here are the observed AGN numbers, without
    correcting for incompleteness.  Further details on each of these
    data sets are in Table~\ref{tab:samples} and
    Section~\ref{sec:sample}.}
  \label{fig:qsos}
\end{figure*}

\begin{table*}
  \caption{AGN samples analysed in this work.}
  \label{tab:samples}
  \begin{tabular}{lcllrS}
    \hline
    Sample& $z$ range$^a$& Survey & Reference & Number & {Area} \\
    & & & & of quasars & {(deg$^2$)} \\
    \hline
    1 & 0.0--2.2 & SDSS DR7 & \citet{2010AJ....139.2360S} & 48664 & 6248.0 \\
    2$^b$ & 0.4--2.2 & 2SLAQ SGP & \citet{2009MNRAS.392...19C} & 2338 & 64.2 \\
    3$^b$ & 0.4--2.2 & --- NGP & \citet{2009MNRAS.392...19C} & 7027 & 127.7 \\
    4 & 2.2--3.5 & BOSS DR9 & \citet{2013ApJ...773...14R} & 23301 & 2236.0 \\
    5 & 3.7--4.7 & SDSS DR7 & \citet{2010AJ....139.2360S} & 1785 & 6248.0 \\
    6 & 3.6--5.2 & NDWFS & \citet{2011ApJ...728L..26G} & 12 & 1.71 \\
    7 & 3.8--5.3 & DLS & \citet{2011ApJ...728L..26G} & 12 & 2.05 \\
    8 & 4.7--5.4 & SDSS+WISE & \citet{2016ApJ...829...33Y} & 99 & 14555.0 \\
    9$^c$ & 4.7--5.5 & SDSS DR7 & \citet{2013ApJ...768..105M} & 103 & 6248.0 \\
    10 & 4.7--5.5 & --- Stripe 82 & \citet{2013ApJ...768..105M} & 62 & 235.0 \\
    11 & 5.7--6.5 & SDSS Main & \citet{2016ApJ...833..222J} & 24 & 11240.0 \\
    12 & 5.7--6.5 & --- Overlap & \citet{2016ApJ...833..222J} & 10 & 4223.0 \\
    13 & 5.7--6.5 & --- Stripe 82 & \citet{2016ApJ...833..222J} & 13 & 277.0 \\
    14 & 5.8--6.6 & CFHQS Deep & \citet{2010AJ....139..906W} & 1 & 4.47 \\
    15 & 5.8--6.6 & --- Very Wide & \citet{2010AJ....139..906W} & 16 & 494.0 \\
    16 & 5.8--6.5 & Subaru High-$z$ Quasar & \citet{2015ApJ...798...28K} & 2 & 6.5 \\
    17$^d$ &4.0--6.5 & CANDELS GOODS-S & \citet{2015AA...578A..83G} & 19 & 0.047 \\
    18$^e$ & 6.5--7.4 & UKIDSS & \citet{2011Natur.474..616M} & 1 & 3370.0 \\
    19$^e$ & 6.5--7.4 & UKIDSS & \citet{2015ApJ...801L..11V} & 1 & 3370.0 \\
    20$^e$ & 6.5--7.4 & ALLWISE+UKIDSS+DECaLS & \citet{2018Natur.553..473B} & 1 & 2500.0 \\
    \hline
  \end{tabular}\\
  \begin{minipage}{14.5cm}
    \textsuperscript{$a$}{Redshift range of the sample or, for small
      samples, approximate redshift range in which the survey is
      sensitive.}\\
    \textsuperscript{$b$}{Restricted to $z<2.2$.}\\
    \textsuperscript{$c$}{Restricted to $M_{1450}>-26.73$ quasars to
      avoid overlap with the \citet{2016ApJ...829...33Y} sample.}\\
    \textsuperscript{$d$}{Used only in Section~\ref{sec:global} and
      Appendix~\ref{sec:conv} due to lack of spectroscopic redshifts
      for majority of the sample.}\\
    \textsuperscript{$e$}{Used only in Section~\ref{sec:global} due to
      roughly estimated selection function.}
 \end{minipage}
\end{table*}

@@

We started with compiling the samples of recent photometric rest-frame UV-optical
quasar surveys. The restriction to UV-optical surveys was mainly
driven by our science goal to characterise the UV luminosity function
of Type~1 quasars. X-ray-selected samples are less suited for this
purpose due to spectroscopic incompleteness and the $\sim 0.4$\,dex
scatter in the conversion from X-ray to UV luminosity
\citep{2010A&A...512A..34L, 2015MNRAS.453.1946G, 2016ApJ...819..154L}
that contributes significantly to the error budget in the UV
luminosity function of X-ray-selected samples unless rest-frame UV
photometry is incorporated \citep{2015AA...578A..83G}.  The individual
surveys and their main characteristics are listed in
Table~\ref{tab:samples}.  Figure~\ref{fig:qsos} presents a redshift
histogram of the contributing surveys.

We included surveys based on a set of simple criteria:
\begin{enumerate}
\item Spectroscopic redshifts for the vast majority of targets.
\item Accurate rest-frame UV-optical CCD photometry.
\item Statistical power (sample size, coverage in $z$ and/or absolute magnitude).
\item A characterised selection function.
\end{enumerate}
As a prerequisite for a joint analysis of the QLF we obtained the
survey selection functions in electronic form, either from the
publication or from the authors. As a reference for future surveys we
publish them here in modified and homogenised form
(Section~\ref{sect:datahom}).

Due to their selection criteria and their statistical power specific
surveys contribute to distinct redshift ranges. At $z<2.2$ we
considered quasars from the SDSS DR7 quasar catalogue
\citep{2010AJ....139.2360S} and the 2SLAQ survey catalogue
\citep{2009MNRAS.392...19C}. We restricted the SDSS DR7 sample to the
48,664 $0.1<z<2.2$ quasars selected with the final SDSS quasar
selection algorithm \citep{2002AJ....123.2945R, 2006AJ....131.2766R}
from a survey area of 6248\,deg$^2$ \citep{2012ApJ...746..169S}. We
adopted the SDSS targeting photometry corrected for Galactic
extinction \citep{2010AJ....139.2360S}. To limit systematic
uncertainties in the correction for host galaxy light
\citep{2009MNRAS.392...19C} we restricted the 2SLAQ sample to 11,816
$g<21.85$ $0.4<z<2.2$ quasars from its spectroscopic survey footprint
near the North Galactic Pole (NGP, 8153 quasars in $127.7$\,deg$^2$)
and the South Galactic Pole (SGP, 3663 quasars in
$64.2$\,deg$^2$). The small sample overlap between SDSS and 2SLAQ (112
quasars) has negligible impact on the QLF evaluation.

@@

The SDSS colours of $5.1<z<5.5$ quasars are similar to those of M and
L dwarf stars, resulting in a low and uncertain selection efficiency
\citep{2013ApJ...768..105M}. WISE mid-infrared selection performs
better \citep{2016ApJ...829...33Y}, but is restricted to the bright
end of the quasar population. Unlike \citet{2013ApJ...768..105M} we
include the 9 uniformly selected $M_{1450}\ge -26.73$ SDSS DR7 quasars
and the 10 SDSS Stripe~82 quasars at $z>5.1$, adopting their low
selection efficiency.  As we will show in Section~3.2, the resulting
QLF is consistent with those at lower and higher redshifts, indicating
that the \citet{2013ApJ...768..105M} selection functions are quite
reliable.

@@

At the highest redshifts $z>6.5$ we considered two quasars found by
\citet{2011Natur.474..616M} and \citet{2015ApJ...801L..11V} in UKIDSS
imaging (3370\,deg$^2$), in addition to the current highest-redshift
quasar at $z=7.54$ selected from a combination of UKIDSS, WISE and
DECaLS \citep[$\sim 2500$\,deg$^2$, ][]{2018Natur.553..473B}. Although
currently only rough estimates exist concerning their selection
functions, these quasars provide constraints on the evolution of the
integrated quasar space density from $z\sim 6$ to $z\sim 7$.  We use
them in our secondary analysis in Section~\ref{sec:global}.  Although
we do not include the highly debated \citet{2015AA...578A..83G} sample
in our main analysis due to its rough selection function and lacking
spectroscopy for 17 of the 22 quasar candidates, we use it in
Section~\ref{sec:global} to constrain the faint end ($M_{1450}>-23$)
of the QLF at $z>4.1$. We restricted the \citet{2015AA...578A..83G}
sample to the 19/22 sources considered in their QLF.

@@

@@

@@

>>>>

\begin{abstract}
  Determinations of the UV luminosity function of AGN at high
  redshifts are important for constraining the AGN contribution to
  reionization and understanding the growth of supermassive black
  holes.  Recent inferences of the luminosity function suffer from
  inconsistencies arising from inhomogeneous selection and mixing of
  AGN data.  We address this problem by constructing a sample of more
  than 80,000 colour-selected AGN from redshift $z=0$ to $7.5$.  While
  this sample is composed of multiple data sets with spectroscopic
  redshifts and completeness estimates, we homogenise these data sets
  to identical cosmologies, intrinsic AGN spectra, and magnitude
  systems.  Using this sample, we derive the AGN UV luminosity
  function from redshifts $z=0$ to $7.5$.  The luminosity function has
  a double power law form at $z<5.5$, and a single power law form at
  higher redshifts.  The break luminosity brightens rapidly with
  redshift and the faint end of the luminosity function steepens.  In
  spite of this steepening, the contribution of AGN to the hydrogen
  photoionization rate at $z\sim 6$ is subdominant, although it can be
  non-negligible if these luminosity functions hold down to
  $M_{1450}=-18$.  Under reasonable assumptions, AGN can reionize
  \HeII\ by redshift $z=3.2$.  At low redshifts ($z<0.5$), AGN can
  produce all of the hydrogen photoionization rate inferred from the
  statistics of \HI\ absorption lines in the IGM.  Our global analysis
  of the luminosity function also reveals important systematic errors
  in the data, particularly at $z=2.2$--$3.5$, which need to be
  addressed in future in order to improve our results.  We make
  various fitting functions, luminosity function analysis codes, and
  homogenized AGN data publicly available.
\end{abstract}

@@

The luminosity function of active galactic nuclei (AGN) and its
evolution over cosmological time scales has been a matter of central
interest of a large body of work over the last four decades
\citep[e.g.,][]{1978A&A....68...17M, 1983ApJ...269..352S,
  1988ApJ...325...92K, 1988MNRAS.235..935B, 1993ApJ...406L..43H,
  1994ApJ...421..412W, 1995AJ....110...68S, 1995AJ....110.2553K,
  1995ApJ...438..623P, 2000MNRAS.317.1014B, 2001AJ....121...54F,
  2004AJ....128..515F, 2006AJ....131.2766R, 2007ApJ...654..731H,
  2009MNRAS.392...19C, 2010AJ....139..906W, 2011ApJ...728L..26G,
  2013ApJ...773...14R, 2013ApJ...768..105M, 2015AA...578A..83G,
  2015ApJ...798...28K, 2016ApJ...829...33Y, 2016ApJ...833..222J,
  2017MNRAS.466.1160M}.  Determination of the AGN luminosity function
constrains models of the build-up of supermassive black holes
\citep{2015MNRAS.452..575S, 2016MNRAS.462..190R}.
%% JFH Some historical papers should be cited for this SMBH buildup point, like
%% for example Soltan. 
Due to the
incidence of supermassive black holes in most galaxies, the tight
scaling relations observed to exist between the mass of these black
holes and properties of their host galaxies
\citep{2013ARA&A..51..511K, 2013ApJ...764..184M}, and the increasing
consensus that AGN activity feeds back on the host galaxy evolution,
the AGN luminosity function also constrains models of galaxy
formation.  Finally, thanks to their relative brightness and high
Lyman-continuum (LyC) escape fractions, the luminosity function of AGN
determines their influence on the temperature and ionization state of
the intergalactic medium (IGM) over large scales,
%% JFH Probably want to say something here about radiation background
%% instead of this ``large scales'' 
possibly even
primarily driving hydrogen and helium reionization.

There is a renewed interest in understanding the evolution of the UV
luminosity function of AGN, triggered by the 19 low-luminosity
($M_{1450}>-22.6$)
%% JFH Probably more sensible to quote a luminosity range rather than just a > 
AGN candidates between redshifts $z=4.1$ and $6.3$
discovered by \citet{2015AA...578A..83G} using a novel X-ray/NIR
selection criterion.  This finding suggested that AGN brighter than
$M_{1450}=-18$ can potentially produce all of the metagalactic
hydrogen photoionization rate inferred from the \lya forest at
$4<z<6$.
%% JFH Probably should say something here about assuming all ionizing photons escape, since these
%% are galaxy like luminosities. Maybe quote L_star galaxy for comparison, which helps with your next
%% point about faint galaxies. 
A significant presence of AGN at high redshift ($z\sim 6$)
and a dominant contribution of AGN to reionization is appealing as the
LyC escape fraction of galaxies is uncertain.  High-redshift galaxies
down to rest-frame UV magnitude $M_\mathrm{UV}=-12.5$ ($L\sim
10^{-3}L^*$) at $z=6$ \citep{2017ApJ...835..113L} and redshifts up to
$z=11.1$ \citep{2016ApJ...819..129O} have now been reported.  But the
escape of LyC photons has been measured in only a few comparatively
bright ($L>0.5L^*$) galaxies at relatively low redshifts ($z < 4$).
The escape fraction in these galaxies is inferred to be typically
about 2--20\% \citep{2010ApJ...725.1011V, 2011ApJ...736...41B,
  2015ApJ...804...17S, 2015ApJ...810..107M, 2016A&A...585A..48G,
  2017MNRAS.468..389J, 2017MNRAS.465..316M}, whereas reionization
requires escape fraction of about 20\% in galaxies as faint as
$M_\mathrm{UV}=-13$ \citep{2016PASA...33...37F, 2015ApJ...802L..19R,
  2016MNRAS.457.4051K}.  An enhanced incidence of high-redshift AGN
may also be consistent with the shallow bright-end slopes of the
$z\sim 7$ UV luminosity function of galaxies relative to a Schechter
function \citep{2012MNRAS.426.2772B, 2014MNRAS.440.2810B,
  2014ApJ...792...76B, 2015MNRAS.452.1817B} and the hard spectra of
these galaxies \citep{2015MNRAS.450.1846S, 2015MNRAS.454.1393S,
  2017MNRAS.464..469S}.  Finally, AGN may also provide a natural
explanation for the large scatter in the \lya opacity between
different quasar sightlines close to redshift $z=6$
\citep{2015MNRAS.447.3402B, 2018arXiv180208177B}.
%% JFH Please cite new paper by eilers et al. here, not yet on archive. 
These opacity
fluctuations extend to substantially larger scales ($\gtrsim 50\,
h^{-1}$cMpc) than expected in galaxies-dominated reionization models
\citep{2015MNRAS.453.2943C}.
%% JFH You probably need a ``but see Davies et al. 201X'' since Fred claims
%% he can explain these fluctuations with galaxies. See also the new paper with George Becker and Fred. 
AGN clustering can result in these
fluctuations naturally if there is a significant contribution
($\gtrsim 50\%$) of AGN to the ionising emissivity at $z=5$--$6$
\citep{2017MNRAS.465.3429C}.

Several counter-arguements against a higher incidence of AGN at high
redshift have also been made. \citet{2017MNRAS.468.4691D} and
\citet{2018MNRAS.473.1416M} pointed out that an AGN population
consistent with the findings of \citet{2015AA...578A..83G} leads to a
much earlier \HeII\ reionization.  For instance, in the model of
\citet{2015ApJ...813L...8M}, He~\textsc{ii} reionization is complete
at $z=4.5$, compared to $z=3$ in the standard scenario
\citep{2012ApJ...746..125H}.  Such early \HeII\ reionization could
result in higher IGM temperatures due to the associated photoheating.
\citet{2017MNRAS.468.4691D} found that the temperature of the IGM at
mean density is twice as much in AGN-dominated reionization models as
the standard models at $z=3.5$--$5$, in conflict with constraints from
the \lya forest.  This inconsistency could be avoided by postulating a
reduced escape fraction of \HeII-ionizing photons in AGN, but it is
difficult to reconcile this with a unit escape fraction of
hydrogen-ionizing photons that is required to explain the Ly$\alpha$
opacity fluctuations.
%% JFH Probably need to clarify for this last point that you mean in a situation
%% where galaxies contribute no LL photons, and QSOs completely dominate. 
Further arguments against AGN-dominated
reionization have been presented by \citet{2016MNRAS.459.2299F} by
analysing the incidence of metal-line absorbers at $z\sim 6$.
\citet{2016MNRAS.459.2299F} find that in their cosmological radiation
hydrodynamical simulations AGN-dominated UV background results in too
many C~\textsc{iv} absorption systems relative to Si~\textsc{iv} and
C~\textsc{ii} at $z\sim 6$.  However, these simulations assume a
$L_\nu\propto\nu^{-1.57}$ AGN SED at extreme UV
\citep{2002ApJ...565..773T}, which somewhat harder than recent
measurements ($L_\nu\propto\nu^{-1.7}$) by
\citet{2015MNRAS.449.4204L}.  \citet{2016MNRAS.459.2299F} also found
that the N(Si~\textsc{iv})/N(C~\textsc{iv}) column density ratio
measurements prefer a harder and more intense $>4$~Ry background than
the standard model \citep{2012ApJ...746..125H}.  Finally, comparing
the \citet{2015AA...578A..83G} sample to X-ray-selected quasar data at
$z=0$--$6$, \citet{2017MNRAS.465.1915R} argued that the faint end of
the AGN UV luminosity function at $z\sim 6$ is probably shallower that
that reported by \citet{2015AA...578A..83G}.
\citet{2017MNRAS.465.1915R} argue that the apparent contradiction with
the results of \citet{2015AA...578A..83G} could be explained by
contamination from the AGN host galaxies for faint AGN.

@@

\begin{figure*}
  \begin{center}
    \includegraphics[width=\textwidth]{qsos.pdf}
    % data.py 
  \end{center}
  \caption{Redshift distribution of the 83,488 AGN considered in this
    analysis.  Shown here are the observed AGN numbers, without
    correcting for incompleteness.  Further details on each of these
    data sets are in Table~\ref{tab:samples} and
    Section~\ref{sec:sample}.}
  \label{fig:qsos}
\end{figure*}

\begin{table*}
  \caption{AGN samples analysed in this work.}
  \label{tab:samples}
  \begin{tabular}{lcllrS}
    \hline
    Sample& $z$ range$^a$& Survey & Reference & Number & {Area} \\
    & & & & of quasars & {(deg$^2$)} \\
    \hline
    1 & 0.0--2.2 & SDSS DR7 & \citet{2010AJ....139.2360S} & 48664 & 6248.0 \\
    2$^b$ & 0.4--2.2 & 2SLAQ SGP & \citet{2009MNRAS.392...19C} & 2338 & 64.2 \\
    3$^b$ & 0.4--2.2 & --- NGP & \citet{2009MNRAS.392...19C} & 7027 & 127.7 \\
    4 & 2.2--3.5 & BOSS DR9 & \citet{2013ApJ...773...14R} & 23301 & 2236.0 \\
    5 & 3.7--4.7 & SDSS DR7 & \citet{2010AJ....139.2360S} & 1785 & 6248.0 \\
    6 & 3.6--5.2 & NDWFS & \citet{2011ApJ...728L..26G} & 12 & 1.71 \\
    7 & 3.8--5.3 & DLS & \citet{2011ApJ...728L..26G} & 12 & 2.05 \\
    8 & 4.7--5.4 & SDSS+WISE & \citet{2016ApJ...829...33Y} & 99 & 14555.0 \\
    9$^c$ & 4.7--5.5 & SDSS DR7 & \citet{2013ApJ...768..105M} & 103 & 6248.0 \\
    10$^c$ & 4.7--5.5 & --- Stripe 82 & \citet{2013ApJ...768..105M} & 59 & 235.0 \\
    11 & 5.7--6.5 & SDSS Main & \citet{2016ApJ...833..222J} & 24 & 11240.0 \\
    12 & 5.7--6.5 & --- Overlap & \citet{2016ApJ...833..222J} & 10 & 4223.0 \\
    13 & 5.7--6.5 & --- Stripe 82 & \citet{2016ApJ...833..222J} & 13 & 277.0 \\
    14 & 5.8--6.6 & CFHQS Deep & \citet{2010AJ....139..906W} & 1 & 4.47 \\
    15 & 5.8--6.6 & --- Very Wide & \citet{2010AJ....139..906W} & 16 & 494.0 \\
    16 & 5.8--6.5 & Subaru High-$z$ Quasar & \citet{2015ApJ...798...28K} & 2 & 6.5 \\
    17$^d$ &4.0--6.5 & CANDELS GOODS-S & \citet{2015AA...578A..83G} & 19 & 0.047 \\
    18$^e$ & 6.5--7.4 & UKIDSS & \citet{2011Natur.474..616M} & 1 & 3370.0 \\
    19$^e$ & 6.5--7.4 & UKIDSS & \citet{2015ApJ...801L..11V} & 1 & 3370.0 \\
    20$^e$ & 6.5--7.4 & ALLWISE+UKIDSS+DECaLS & \citet{2018Natur.553..473B} & 1 & 2500.0 \\
    \hline
  \end{tabular}\\
  \begin{minipage}{14.5cm}
    \textsuperscript{$a$}{Redshift range of the sample or, for small
      samples, approximate redshift range in which the survey is
      sensitive.}\\
    \textsuperscript{$b$}{Restricted to $z<2.2$.}\\
    \textsuperscript{$c$}{Restricted to $M_{1450}>-26.73$ quasars to
      avoid overlap with the \citet{2016ApJ...829...33Y} sample.}\\
    \textsuperscript{$d$}{Used only in Section~\ref{sec:global} and
      Appendix~\ref{sec:conv} due to lack of spectroscopic redshifts
      for majority of the sample.}\\
    \textsuperscript{$e$}{Used only in Section~\ref{sec:global} due to
      roughly estimated selection function.}
 \end{minipage}
\end{table*}

@@

We started with
%% JFH we started with --> we started by
compiling the samples of recent photometric rest-frame UV-optical
quasar surveys. The restriction to UV-optical surveys was mainly
driven by our science goal to characterise the UV luminosity function
of Type~1 quasars. X-ray-selected samples are less suited for this
purpose due to spectroscopic incompleteness and the $\sim 0.4$\,dex
scatter in the conversion from X-ray to UV luminosity
\citep{2010A&A...512A..34L, 2015MNRAS.453.1946G, 2016ApJ...819..154L}
that contributes significantly to the error budget in the UV
luminosity function of X-ray-selected samples unless rest-frame UV
photometry is incorporated \citep{2015AA...578A..83G}.  The individual
surveys and their main characteristics are listed in
Table~\ref{tab:samples}.  Figure~\ref{fig:qsos} presents a redshift
histogram of the contributing surveys.

We included surveys based on a set of simple criteria:
\begin{enumerate}
\item Spectroscopic redshifts for the vast majority of targets.
  %% JFH Did we actually use any photometric redshifts? If not, I would omit this ``Vast majority'' and make
  %% it more definite. 
\item Accurate rest-frame UV-optical CCD photometry.
\item Statistical power (sample size, coverage in $z$ and/or absolute magnitude).
\item A characterised selection function.
  %% JFH well characterized
\end{enumerate}
As a prerequisite for a joint analysis of the QLF we obtained the
survey selection functions in electronic form, either from the
publication or
%% JFH by request 
from the authors. As a reference for future surveys we
publish them here
%% JFH rather than publish them here say ``we make them electronically available here''. Cite an appendix where
%% you do this. 
in modified and homogenised form
(Section~\ref{sect:datahom}).

Due to their selection criteria and their statistical power specific
surveys contribute to distinct redshift ranges. At $z<2.2$ we
considered quasars from the SDSS DR7 quasar catalogue
\citep{2010AJ....139.2360S} and the 2SLAQ survey catalogue
\citep{2009MNRAS.392...19C}. We restricted the SDSS DR7 sample to the
48,664 $0.1<z<2.2$ quasars selected with the final SDSS quasar
selection algorithm \citep{2002AJ....123.2945R, 2006AJ....131.2766R}
from a survey area of 6248\,deg$^2$ \citep{2012ApJ...746..169S}. We
adopted the SDSS targeting photometry corrected for Galactic
extinction \citep{2010AJ....139.2360S}. To limit systematic
uncertainties in the correction for host galaxy light
%% JFH What is Croom the reference for here, host galaxy contamination or the 2SLAQ sample. You basically
%% need a reference for 2SLAQ and it appears in the wrong place here. 
\citep{2009MNRAS.392...19C} we restricted the 2SLAQ sample to 11,816
$g<21.85$ $0.4<z<2.2$ quasars from its spectroscopic survey footprint
near the North Galactic Pole (NGP, 8153 quasars in $127.7$\,deg$^2$)
and the South Galactic Pole (SGP, 3663 quasars in
$64.2$\,deg$^2$). The small sample overlap between SDSS and 2SLAQ (112
quasars) has negligible impact on the QLF evaluation.
%% JFH Maybe state what the 2SLAQ survey goal was re quasars.

@@

The SDSS colours of $5.1<z<5.5$ quasars are similar to those of M and
L dwarf stars, resulting in a low and uncertain selection efficiency
\citep{2013ApJ...768..105M}. WISE mid-infrared selection performs
better \citep{2016ApJ...829...33Y}, but is restricted to the bright
end of the quasar population. Unlike \citet{2013ApJ...768..105M} we
include the 9 uniformly selected $M_{1450}\ge -26.73$ SDSS DR7 quasars
and the 10 SDSS Stripe~82 quasars at $z>5.1$, adopting their low
selection efficiency.
%% JFH efficiency and completeness are two different things and you seem to be conflating
%% the two concepts here. 
As we will show in Section~3.2, the resulting
QLF is consistent with those at lower and higher redshifts, indicating
that the \citet{2013ApJ...768..105M} selection functions are quite
reliable.

@@

At the highest redshifts $z>6.5$ we considered two quasars found by
\citet{2011Natur.474..616M} and \citet{2015ApJ...801L..11V} in UKIDSS
imaging (3370\,deg$^2$)
%% JFH quote the individual redshifst of the mortlock and venemans objects
, in addition to the current highest-redshift
quasar at $z=7.54$ selected from a combination of UKIDSS, WISE and
DECaLS \citep[$\sim 2500$\,deg$^2$, ][]{2018Natur.553..473B}. Although
currently only rough estimates exist concerning their selection
functions, these quasars provide constraints on the evolution of the
integrated quasar space density from $z\sim 6$ to $z\sim 7$.  We use
them in our secondary analysis in Section~\ref{sec:global}.  Although
we do not include the highly debated \citet{2015AA...578A..83G} sample
in our main analysis due to its rough selection function and lacking
%% JFH lacking --> lack of 
spectroscopy for 17 of the 22 quasar candidates, we use it in
Section~\ref{sec:global} to constrain the faint end ($M_{1450}>-23$)
of the QLF at $z>4.1$. We restricted the \citet{2015AA...578A..83G}
sample to the 19/22 sources considered in their QLF.

@@

%% JFH This section is long tedious and hard to follow. Perhaps we
%% should put a brief description here, relegate these details to the
%% appendix, and in the appendix, structure this section by
%% sub-sections for each individual survey.

@@

%% JFH Let's make all these conversion factors and codes to do K-correction
%% publicly avaialble in an appendix.

@@

%% JFH We need a better roadmap here for where this section is going. 1) Binned estimates 2) individual
%% fits in redshift bins. 3) examining redshift evolution. 4) attempting a joint fit.
